---
title: "Fuzzy Matching HOH_IDs"
author: "Leandro Ledesma"
date: "2024-02-25"
output: html_document
---

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(kableExtra)
library(openxlsx) # To save excel files with multiple tabs
library(fuzzyjoin)
```

### Load in the data 

We will be loading in data that most likely has typos in the name. 


```{r load in the data with typos}
# Set the working directory
setwd("C:/Users/lledesma.TIMES/Documents/KBB_new_2/1_screener/final_data")

# Load in the data
potential.HOH_ID.errors <- read_excel("Potential_Household_errors.xlsx", sheet = "error_household")
correct.HOH_IDs <- read_excel("Potential_Household_errors.xlsx", sheet = "correct_household")

```

### Fuzzy matching (Part 1)

We will now take the typo ID's and try to match them, mathematically, to the closest ID that was correctly typed. I have set the code to only show the matches that have a distance value of less than 4. These have the highest chance of being correct typo to real HOH_ID matches. Additionally, use the variables 'Date of Evaluation' and 'Village' to increase your confidence that these HOH_IDs are the same. Again, just because the code matched them, doesn't mean they should be matched, but there is a highly probability that these two HOH_IDs are the same. 

```{r fuzzy matching}
# Data cleaning
potential.HOH_ID.errors <- potential.HOH_ID.errors %>%
  select(HOH_ID, Date_of_Evaluation, Name_of_the_Village, Evaluator_ID)

correct.HOH_IDs <- correct.HOH_IDs %>%
  select(HOH_ID, Date_of_Evaluation, Name_of_the_Village,Evaluator_ID)

# Run the fuzzy matching
fuzzy.matching <- stringdist_join(potential.HOH_ID.errors, correct.HOH_IDs, 
                by= "HOH_ID", #match based on ID
                mode='left', #use left join
                method = "osa", #use jw distance metric
                max_dist=99, 
                distance_col='dist') %>%
  group_by(HOH_ID.x) %>%
  slice_min(order_by=dist, n=1) %>%
  data.frame() %>%
  rename(`Typo HOH_ID` = HOH_ID.x, 
         `Correct HOH_ID` = HOH_ID.y) %>%
  arrange(dist) %>%
  filter(dist < 4) %>%
  unique()

paste("From the Fuzzy Matching code, there are potentially",nrow(fuzzy.matching),"typo IDs that were matched to real HOH_IDs")

# Save only the unique matches
fuzzy.matching <- unique(fuzzy.matching)

# Print out the options
fuzzy.matching %>%
  arrange(desc(Date_of_Evaluation.x)) %>%
  kbl() %>%
  kable_paper(bootstrap_options = "striped", full_width = F)


# Remove the typo HOH_IDs that were matched to a correct HOH_ID from the 'potential.HOH_ID.errors' data frame
potential.HOH_ID.errors2 <- potential.HOH_ID.errors %>%
  filter(!(HOH_ID %in% fuzzy.matching$`Typo HOH_ID`))

```


### Fuzzy matching (Part 2)

There are, however, other cases where there are only two instances where an HOH_ID is typed, and neither version matches each other. For these cases, we need to one by one match a typo HOH_ID to the rest of the typo HOH_IDs and see if we can potentially find some matches here. 


```{r fuzzy matching part 2}
paste("There are",nrow(potential.HOH_ID.errors2),"typo HOH_IDs that still need to be matched")

# Create a function to take on HOH_ID by one and then match to the rest in the data frame
potential.matches.list <- list()

for(ii in 1:nrow(potential.HOH_ID.errors2)) {

typo.ID.length <- nrow(potential.HOH_ID.errors2)
current.row <- ii
not.current.row <- setdiff(1:typo.ID.length, current.row)

current.row.df <- potential.HOH_ID.errors2[current.row,]
not.current.row.df <- potential.HOH_ID.errors2[not.current.row,]


potential.matches.list[[ii]] <- stringdist_join(current.row.df, not.current.row.df, 
                by= "HOH_ID", #match based on ID
                mode='left', #use left join
                method = "osa", #use jw distance metric
                max_dist=99, 
                distance_col='dist') %>%
  rename(`Typo HOH_ID 1` = HOH_ID.x, 
         `Type HOH_ID 2` = HOH_ID.y) %>%
  arrange(dist) %>%
  filter(dist < 4) %>%
  unique()

}

# Print the new potential matches 
fuzzy.matching2 <- potential.matches.list %>%
  do.call(rbind,.)
  
fuzzy.matching2 %>%
    arrange(desc(Date_of_Evaluation.x)) %>%
    kbl() %>%
    kable_paper(bootstrap_options = "striped", full_width = F)

# Remove the typo HOH_IDs that were matched to another typo HOH_ID from the 'potential.HOH_ID.errors' data frame
potential.HOH_ID.errors3 <- potential.HOH_ID.errors2 %>%
  filter(!(HOH_ID %in% fuzzy.matching2$`Typo HOH_ID 1`))

```


### ID's that could not be matched at all

The ID's down below could not be matched to any existing HOH_ID's nor to each other. This could be the case for the following reasons:

1) There was a mistake with designating HOH information and respondent information (such as names or birthdates)

2) The typos are so severe that they cannot be matched (information was typed in very incorrectly)

3) Only one HOH_ID was asked for a child- which does not make sense

4) The matching ID was either lost, not saved, or not updated to KoboToolBox website yet. 


```{r IDs that could not be matched at all}
potential.HOH_ID.errors3 %>%
  unique() %>%
  arrange(desc(Date_of_Evaluation)) %>%
  kbl() %>%
    kable_paper(bootstrap_options = "striped", full_width = F)

```



