---
title: "Calculating a correlation by hand"
author: "Leandro Ledesma"
date: "2024-02-01"
output: html_document
---
### Our data

We will be using the following data from the Data Sets of the 'Statistical Methods for Psychology by David C. Howell' website. This data set contains the variables Height and Weight. We will be calculating by hand the degree of relationship between these two variables.

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(message = FALSE)
```

### Load in the data manipulation packages

```{r load in packages, warning = FALSE}
library(tidyverse)
library(ggplot2)

```

### Load in the data

```{r load the data}
# Load in the data
data <- read.delim("https://www.uvm.edu/~statdhtx/methods8/DataFiles/Ex9-31.dat")

# Explore the data
dim(data)
str(data)

```

### Plot the data


```{r plot the data}
# Create a scatterplot
data %>%
  ggplot(aes(x = Height, y = Weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(title = "Scatterplot of Height and Weight",
       x = "Height",
       y = "Weight") +
  theme(plot.title = element_text(size = 15,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))


```

