---
title: "Adding Demographic Information"
output: html_document
---

### Universal block code settings

```{r universal block code setting}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
```

### Load in the data manipulation packages first

```{r load in packages, message= FALSE, warning = FALSE}
# Load in the proper packages
library(tidyverse)
library(readr)
library(readxl)
library(kableExtra)
```

### Last day this code was ran

```{r last day the code was ran}

paste("This code was last ran on:", Sys.Date())

```

### Load in the data

We will be loading demographic information for the ONR study and then merging it with our GnG and Nback CSVs. Demographic information is saved in a screener that is used to assess whether participants are eligible to participate in the study. Unfortunately this screener does not contain ID numbers for the participants. However, we can get around this by merging it with the SharePoint ID's since they contain matching email addressed- this is what we will be doing below.

Message: Had to drop ID's 14 and 90 because their inclusion led to weird merger between the two datasets below. Might be from having multiple screening emails. 

```{r load in demographic information, warning = FALSE}
# Set working directory
setwd("C:/Users/lledesma.TIMES/Documents/ONR/Demographics")

# Load in the following data
PilotIDs <- read.csv("PilotIDs.csv")
Screener.Ano <- read.csv("Anonymous.Screener.csv")
Screener.SONA <- read.csv("SONA.Screener.csv")
Screener.ROTC <- read.csv("ROTC.Screener.csv")

# Clean the pilot data by dropping NA's from the ID variable
PilotIDs <- PilotIDs %>%
  drop_na(ParticipantID)

# Check the dimensions of the screeners
dim(Screener.Ano)
dim(Screener.SONA)
dim(Screener.ROTC)

# Remove the first two rows of the Screeners (They are meaningless)
Screener.Ano <- Screener.Ano[3:nrow(Screener.Ano),]
Screener.SONA <- Screener.SONA[3:nrow(Screener.SONA),]
Screener.ROTC <- Screener.ROTC[3:nrow(Screener.ROTC),]


# Select the variables of interest from the screeners
Screener.Ano <- Screener.Ano %>%
  select(email = Q94,
         race = RE1,
         ethnicity = RE2,
         Sex,
         Age = AgeYears) %>%
  data.frame() %>%
  unique()

Screener.SONA <- Screener.SONA %>%
  select(email = Q94,
         race = RE1,
         ethnicity = RE2,
         Sex,
         Age = AgeYears) %>%
  data.frame() %>%
  unique()

Screener.ROTC <- Screener.ROTC %>%
  select(email = RecipientEmail,
         race = RE1,
         ethnicity = RE2,
         Sex,
         Age = AgeYears) %>%
  data.frame() %>%
  unique()
  

# Bind all of the Screeners into one dataset
Screener <- Screener.Ano %>%
  rbind(Screener.SONA) %>%
  rbind(Screener.ROTC)


# Select the variables of interest for PilotIDs
PilotIDs <- PilotIDs %>%
  select(email = Email,
         ID = ParticipantID) %>%
  data.frame()

# Drop ID's 90 and 14 because they come out weird in the dataset
PilotIDs <- PilotIDs %>%
  filter(!(ID %in% c("14","90")))


# Merge these two datasets to obtain the demographics for each ID
demographics <- Screener %>%
  merge(PilotIDs, by = "email")

# Drop the row if the ID column contains an NA
demographics <- demographics %>%
  drop_na(ID)

# Some extra data cleaning
demographics <- demographics %>%
  select(ID, everything()) %>%
  arrange(ID)

```
### Data quality inspection of ID's with demographics

There are some duplicates in our data which were caused by multiple screeners filled out by recruited participants. Some of these cases are easy to fix since they have a duplicate with NA's. Thus, those NA rows can be dropped. For the others, these duplicates in data are caused by taking the screeners in between birthdays. Therefore, we have some participants that took the screener at one age and then retook it after their birthdate, creating duplicates. For these cases we will choose the row with the oldest age. Subject 121 reported different Sexes for their screeners. After checking additional demographic information their wrongly chosen sex was dropped, removing the duplicate. Similarly, subject 190 reported being African American in one screener and More than one race in the second, thus we kept the later for their demographic information. 


```{r unable to get demographics}
# Check for duplicates in the data
demographics %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  kbl() %>%
  kable_paper(full_width = F)

# Some of these duplicates are due to not completing a screener
# We can eliminate these by droppin rows with NA's
# Replace any blank cells with NA's
demographics[demographics == ""] <- NA

# Check for any duplicates in the data or missing data
demographics %>%
  filter(!complete.cases(.)) %>%
  kbl() %>%
  kable_paper(full_width = F)


# Only keep completed cases 
demographics <- demographics %>%
  filter(complete.cases(.))

# Check for duplicates in the data
demographics %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  kbl() %>%
  kable_paper(full_width = F)


# Save these duplicate ID's
duplicate.IDs <- demographics %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  select(ID) %>%
  unlist()

# Duplicate ID's data frame
duplicate.IDs.df <- demographics %>%
  filter(ID %in% duplicate.IDs)


# Remove the duplicate ID's from the demographics data frame
demographics <- demographics %>%
  filter(!(ID %in% duplicate.IDs))

# Create a for loop to keep the rows for each participant with the oldest age
duplicate.IDs.age.corrected <- list()
duplicate.IDs <- unique(duplicate.IDs.df$ID)

for(ii in 1:length(duplicate.IDs)) {

  # Obtain the duplicate rows of the current ID
  current.ID <- duplicate.IDs[ii]
  
  # Calculate the maximum age for this ID
  max.age <- duplicate.IDs.df %>% filter(ID == current.ID) %>% select(Age) %>% unlist() %>% as.numeric() %>% max()
  
  # Save the row of the max age
  duplicate.IDs.age.corrected[[ii]] <- duplicate.IDs.df %>% filter(ID == current.ID & Age == max.age)
}

# Take the corrected for duplicate age IDs and turn them into a df
unlisted.duplicate.IDs.age.corrected  <- duplicate.IDs.age.corrected %>% do.call(rbind,.)

# Manual corrections
corrected.demographics <- unlisted.duplicate.IDs.age.corrected %>%
  filter(!(ID == 121 & Sex == "Male")) %>%
  filter(!(ID == 190 & race == "Black or African American"))

# Introduce the corrected demographics back to the main demographic dataset
demographics <- demographics %>%
  rbind(corrected.demographics)

# Quality control to see if there are any remaining duplicates or missing data
demographics %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  kbl() %>%
  kable_paper(full_width = F)


demographics %>%
  filter(!complete.cases(.)) %>%
  kbl() %>%
  kable_paper(full_width = F)


# Number of unique ID's in this data frame
length(unique(demographics$ID))

```

### Adding demographics manually

Unfortunately, if participants enrolled reported an email for the screener that was mistyped or decided to participate in the study with a different email- then their demographic information would have been missed with the code above. Thus for these cases, their demographic information was added manually.  


```{r manual demographics, warning= FALSE}
# ID's that need demographic data
setdiff(PilotIDs$ID, demographics$ID) %>% sort()

# Set working directory
setwd("~/ONR/Demographics")

# Load in the manual demographic sheet
manual.demographics <- read_excel("demographics.manual.xlsx")

# no demographic information avaiable
paste("Warning, no demographic information could be found for the following ID:", manual.demographics$ID[is.na(manual.demographics$Age)],sep=" ")


# Add manual demographics to the rest of our demographics dataset
manual.demographics$email <- NA
manual.demographics$race <- NA
manual.demographics$ethnicity <- NA

manual.demographics <- manual.demographics %>% select(ID, email, race, ethnicity, Sex, Age)

demographics <- demographics %>%
  rbind(manual.demographics)


```
### Current demographic information numbers

```{r count demographic info numbers}
paste("We have demographic information for:", length(unique(demographics$ID)),"participants out of", length(unique(PilotIDs$ID)), sep=" ")

```

### Save demographic information

```{r save demographic information, warning = FALSE}
# Set save working directory
setwd("C:/Users/lledesma.TIMES/Documents/ONR/Demographics")

# Save the data as a CSV
write.csv(x = demographics, file = "complete.demographics.csv")

```
