---
title: "Analyzing Vigilance Decrement in the GnG"
author: "Leandro Ledesma"
date: "2024-04-04"
output: html_document
---

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readxl)
library(lmtest) # bptest() function for homoscedasticity
library(kableExtra)
library(boot)

```

### Load in the data

As mentioned above, we are only interested in performance on Day 1 for the 'Colors and Shape' block session for the GnG and the forward arrows no distractor condition in the Nback. Thus, these data will remain for the rest of the analysis. 

```{r load in the data, warning = FALSE}
# Set working directory
setwd("C:/Users/lledesma.TIMES/Documents/ONR/GnG_and_Nback_behavioral_data_preprocessing/Processed Data")

# Load in the Nback data (Reacted)
GnG.ordered <- read.csv("GnG.conditions.ordered.csv")

# Data cleaning- we are only interested in Day 1
GnG.day1 <- GnG.ordered %>% filter(Day.Type == "Day 1") 

# Adding demographic data
setwd("C:/Users/lledesma.TIMES/Documents/ONR/Demographics")

# Save the data as a CSV
demographics <- read.csv("complete.demographics.csv")

# Clean demographics
demographics <- demographics %>%
  select(-c(X, email))

```



### Graph the data across time

```{r graphing the data}
GnG.day1 %>%
  ggplot(aes(x=Order, 
             y = trial.mean.order.correct, 
             color = Trial.Type)) +
  stat_summary(fun = mean,
               geom = "point") +
  stat_summary(fun = mean,
               geom = "line") +
  stat_summary(fun.data = mean_cl_boot,
               geom ="errorbar",
               width = .2)

data.frame(rbind(table(GnG.day1$Order, GnG.day1$BlockName)))

```


### View the distribution of each group order

```{r graphing each group order}

GnG.day1 %>%
  filter(Order == 5) %>%
  ggplot(aes(x = trial.mean.order.correct)) +
  geom_histogram() +
  facet_wrap(~Trial.Type)



```

### Analyzing the data

```{r analyzing the data with a robust ANOVA type}
GnG.day1$Day.Type <- factor(GnG.day1$Day.Type)
GnG.day1$Order <- factor(GnG.day1$Order)
GnG.day1$BlockName <- factor(GnG.day1$BlockName)
GnG.day1$UserID <- factor(GnG.day1$UserID)

library(rstatix)

GnG.day1.Go <- GnG.day1 %>% filter(Trial.Type == "Go")
GnG.day1.NoGo <- GnG.day1 %>% filter(Trial.Type == "No-Go")

anova_test(data = GnG.day1.NoGo, 
           dv = trial.mean.order.correct, 
           wid = UserID, 
           within = Order)

summary(x)


library(lme4)
library(lmerTest)

# Make the first order the reference one
GnG.day1 <- within(GnG.day1, Order <- relevel(Order, ref = 2))
GnG.day1$Order <- relevel(GnG.day1$Order, "1")


model <- lmer(trial.mean.order.correct ~ Order + (1 | UserID),
              data = GnG.day1)

summary(model)
```
