---
title: "CFIT Scoring"
author: "Leandro Ledesma"
date: "2024-02-14"
output: html_document
---


The CFIT stands for the Culturally Fair Intelligence Tests developed by Raymond Cattell. This assessment contains four types of non-verbal IQ tests. We are interested in Scale 2 of this test, which is the medium difficulty version. Additionally Scale 2 contains Form A and Form B, both forms contain items that correspond to one of the four tests. We collected data strictly from Form B. Unfortunately, there is no standardized way to obtain IQ scores from performance on the CFIT Scale 2 Form B, thus we will only be reporting raw scores for each subject below. 

Raw scores are calculated by taking the sum of correct responses from the four subtests. Meaning there will be one overall score. 

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readxl)
library(kableExtra)

```

### Load in the data

```{r load in the data, warning = FALSE}
# Set working directory
setwd("~/ONR/IQ_data")

# Load in the data
CFIT <- read.csv("ONR CFIT_February 14, 2024_15.14.csv")

```

### Data cleaning

```{r data cleaining}
# Remove unnecessary rows 
# Remove the first seven rows they are meaningless
CFIT <- CFIT[8:nrow(CFIT),]

# Keep variables that we are interested in
CFIT.names <- names(CFIT)
series.names <- CFIT.names[grepl(pattern = "T1Q" , x = CFIT.names)]
classification.names <- CFIT.names[grepl(pattern = "T2Q" , x = CFIT.names)]
matrices.names <- CFIT.names[grepl(pattern = "T3Q" , x = CFIT.names)]
condition.names <- CFIT.names[grepl(pattern = "T4Q" , x = CFIT.names)]

# The number of items in order should be 12, 14, 12, and 8
length(series.names)
length(classification.names)
length(matrices.names)
length(condition.names)

# Extract these variable names from our original CFIT
variables <- c("ID", series.names, classification.names, condition.names)

# Save this new cleaned data set
CFIT <- CFIT[,variables]
```


### Quality control

First let's remove any NA's from our ID column since this is meaningless information. Then let's check for any duplicates in our data which is from tester error. Below are some ID's that have more than two scores for IQ. To fix these scores, we would have to look at the dates of completion between these ID's and neighboring ID's, which may take a while. For now these ID's will be removed from the analysis. 

```{r quality control}
# Convert any blank information in the cells into NA
CFIT[CFIT == ""] <- NA

# Dropping NA's in the ID column
CFIT <- CFIT %>%
  drop_na(ID)

# Look for duplicates in the data
CFIT %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  kbl() %>%
  kable_paper(full_width = F)

# Duplicate IDs
duplicate.IDs <- CFIT %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  select(ID) %>%
  unlist()

# Mentioned how many ID's will be removed
paste("We will be dropping",length(duplicate.IDs),"IDs from our analysis.")

# Drop duplicate IDs
CFIT <- CFIT %>%
  filter(!(ID %in% duplicate.IDs))

# Check for any remaining duplicate IDs
CFIT %>%
  group_by(ID) %>%
  count() %>%
  filter(n > 1) %>%
  kbl() %>%
  kable_paper(full_width = F)

```



### CFIT Scoring

We are interested in obtain one raw score per person, thus we will be matching every response to that of the response key. Trials where the answer and response key match will be given a score of 1 and non-matching trials will be given a 0. All points will then be added together to produce our raw score. 

```{r CFIT scoring, warning = FALSE}
# Convert the data to long
CFIT.long <- CFIT %>%
  pivot_longer(-ID)

# Set working directory
setwd("~/ONR/IQ_data")

# Load in the answer key 
answer.key <- read_excel("CFIT Raw Score Key.xlsx")

# Create a for loop to score each partici

```




