---
title: "ONR Day Labeling"
author: "Leandro Ledesma"
date: "2024-02-05"
output: html_document
---

### Universal block code settings

```{r universal block code setting}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
```

### Load in the data manipulation packages first

```{r load in packages, message= FALSE, warning = FALSE}
# Load in packages
library(tidyverse)
library(readxl)
library(gdata) # Use startsWith function
library(base) #subset strings
library(scales)
library(rio)
library(purrr)
```


### Create ID's with leading zeroes

```{r first block}
# Create an empty character vector to store ID numbers
ID.numbers <- character(999)

# Use sprintf to format the values with leading zeros so it creates a vector with zeroes in fron of each number smaller than 100
for (i in 1:999) {
  ID.numbers[i] <- sprintf("%03d", i)
}

# Check to see if the zeroes were added
head(ID.numbers,99)

```


### Load in the data

These next few sections are a bit complex. ONR behavioral data is saved locally on two separate computers (UH-ONR and UH-ONR2). This is because we are using two rooms for data collection, each with their own VR systems. Every day around 3:00 AM, data saved locally on these computers are transfered to a secure server where they are saved in separate folders. Inside these directories, there are folders with ID names (ex: 001 or 015) that contain CSVs representing performance during the VR task. Each CSV corresponds to a GnG or an Nback task completed on one of the four testing days. Additionally these CSV files are named weirdly (foreshadowing a problem that will have to be fixed). 

This first chunk of code is going into both directories on the server and recording the pathways of participants that have behavioral data saved. 

```{r obtaining pathways with behavioral data for each subject, warning = FALSE}
# set working directory for ONR computer 1
setwd("M:/VR_CSVs/UH-ONR")

# let's get the names of all the files in the folder
file.names.UH_ONR <- list.files()

# set working directory for ONR computer 1
setwd("M:/VR_CSVs/UH-ONR2")

# let's get the names of all the files in the folder
file.names.UH_ONR2 <- list.files()

# Obtain the intersected IDs from both computers
intersect.IDs.UH_ONR <- intersect(file.names.UH_ONR,ID.numbers)
intersect.IDs.UH_ONR2 <- intersect(file.names.UH_ONR2,ID.numbers)

# Create two pathways using the intersected IDs
ID.folder.UH_ONR <- paste("M:/VR_CSVs/UH-ONR/",intersect.IDs.UH_ONR,"/", sep="")
ID.folder.UH_ONR2 <- paste("M:/VR_CSVs/UH-ONR2/",intersect.IDs.UH_ONR2,"/", sep="")

```

This next chunk of code is using the pathways from above to record the names and pathways of each CSV file in Computer 1 and saving them as a data frame. 

```{r obtain pathway csvs and computer data frame for computer 1}

# Create a for loop to go into each pathway of a subject's ID folder and extract the names of CSV's present there
# Do this for both pathways
# Computer 1
CSVs.UH_ONR <- list()

for(ii in 1:length(ID.folder.UH_ONR)) {
  #ii = 1
  # Changes the working directory to each pathway
  setwd(ID.folder.UH_ONR[[ii]])
  
  # Obtain the CSVs for the folder
  CSVs.current <- list.files()
  
  # Create a vector for the current ID and computer number (Repeat it by the number of CSVs present)
  current.ID <- rep(intersect.IDs.UH_ONR[ii], length(CSVs.current))
  current.pathway <- rep(pathway.UH_ONR[ii], length(CSVs.current))
  current.computer <- rep("Computer 1", length(CSVs.current))
  
  # Create a data frame including the ID, the pathway, CSV, and the computer number
  UH_ONR.df <- data.frame(ID = current.ID,
                          Pathway = current.pathway,
                          CSVs = CSVs.current,
                          Computer = current.computer) 
  
  # Extracts all the names of the CSVs located in that pathway + it's pathway
  CSVs.UH_ONR[[ii]] <- UH_ONR.df
}

CSVs.UH_ONR <- CSVs.UH_ONR %>%
  do.call(rbind,.) %>%
  data.frame()

```


This next chunk of code does the same as above except for the files in computer 2.  

```{r obtain pathway csvs and computer data frame for computer 2}

# Create a for loop to go into each pathway of a subject's ID folder and extract the names of CSV's present there
# Do this for both pathways
# Computer 2
CSVs.UH_ONR2 <- list()

for(ii in 1:length(ID.folder.UH_ONR2)) {
  #ii = 1
  # Changes the working directory to each pathway
  setwd(ID.folder.UH_ONR2[[ii]])
  
  # Obtain the CSVs for the folder
  CSVs.current <- list.files()
  
  # Create a vector for the current ID and computer number (Repeat it by the number of CSVs present)
  current.ID <- rep(intersect.IDs.UH_ONR2[ii], length(CSVs.current))
  current.pathway <- rep(pathway.UH_ONR2[ii], length(CSVs.current))
  current.computer <- rep("Computer 2", length(CSVs.current))
  
  # Create a data frame including the ID, the pathway, CSV, and the computer number
  UH_ONR.df <- data.frame(ID = current.ID,
                          Pathway = current.pathway,
                          CSVs = CSVs.current,
                          Computer = current.computer) 
  
  # Extracts all the names of the CSVs located in that pathway + it's pathway
  CSVs.UH_ONR2[[ii]] <- UH_ONR.df
}

CSVs.UH_ONR2 <- CSVs.UH_ONR2 %>%
  do.call(rbind,.) %>%
  data.frame()

```




# Merge pathways from UHONR1 and UHONR2
# Identify which pathways have redundancies from the second computer and remove them
ONR1_ID_and_CSVs <- sapply(ONR1_All_pathways_to_be_read, function(x) substring(x, 19, nchar(x)))
names(ONR1_ID_and_CSVs) <- NULL
ONR1_ID_and_CSVs_df <- data.frame(Pathway = rep("M:/VR_CSVs/UH-ONR/",length(ONR1_ID_and_CSVs)),
                                  CSVs = ONR1_ID_and_CSVs)


ONR2_ID_and_CSVs <- sapply(ONR2_All_pathways_to_be_read, function(x) substring(x, 20, nchar(x)))
names(ONR2_ID_and_CSVs) <- NULL
ONR2_ID_and_CSVs_df <- data.frame(Pathway = rep("M:/VR_CSVs/UH-ONR2/",length(ONR2_ID_and_CSVs)),
                                  CSVs = ONR2_ID_and_CSVs)

# These are the CSVs that are NOT gonna be read from UHONR2 because that's redundant
Intersection_CSVs <- intersect(ONR1_ID_and_CSVs_df$CSVs,ONR2_ID_and_CSVs_df$CSVs)

# Bring in the cool code
`%nin%` = Negate(`%in%`)

ONR2_ID_and_CSVs_df <- ONR2_ID_and_CSVs_df %>%
  filter(CSVs %nin% Intersection_CSVs)

# Stack the two data frames then combine the variables into one
All_pathways_no_redundancies <- rbind(ONR1_ID_and_CSVs_df,ONR2_ID_and_CSVs_df)
Load_All_Pathways_no_redundancies <- paste(All_pathways_no_redundancies$Pathway,All_pathways_no_redundancies$CSVs,sep="")


# Identify which file came from which computer
VR_Room <- substr(Load_All_Pathways_no_redundancies,15,18)
IDs <- substr(Load_All_Pathways_no_redundancies,19,22)

# A little data cleaning
VR_Room <- gsub("/","", VR_Room)
IDs <- gsub("/","",IDs)

UserID_VR_Room <- data.frame(UserID = IDs,
                             VR_Room)


# Read in all  files
All_CSV_Files <- list()

for(ii in 1:length(Load_All_Pathways_no_redundancies)) {
  
  All_CSV_Files[[ii]] <- read_csv(Load_All_Pathways_no_redundancies[ii],col_names = TRUE, show_col_types = FALSE)
  
}


# Use the map function to create a tibble were all data frames are embedded as rows
All_Data <- data.frame(file_num = 1:length(All_CSV_Files)) %>% as_tibble()
All_Data$Tibbles <- map(All_CSV_Files, as_tibble)

# Create for loops that go into each one of these data frames and retrieve information of interest only from the first row
# This will include UserID and StartDate

UserID <- list()
StartDate <- list()
Task_Type_list <- list()

for(ii in 1:nrow(All_Data)) {

Current_dataset <- All_Data$Tibbles[[ii]]
UserID[ii] <- Current_dataset$UserID[1]
Date_long <- Current_dataset$StartDate[1]
StartDate[ii] <- substr(Date_long,1,10)

Task_type <- Current_dataset$BlockID[1]

if(substr(Task_type,1,1) == "G") {
  Task_Type_list[ii] <- "Go/NoGo"

} else if(substr(Task_type,1,1) == "N") {
  Task_Type_list[ii] <- "N-back"
  
}

}

All_Data$UserID <- unlist(UserID)
All_Data$StartDate <- unlist(StartDate)
All_Data$Task_Type <- unlist(Task_Type_list)
All_Data$User_ID_check <-UserID_VR_Room$UserID
All_Data$VR_Room <- UserID_VR_Room$VR_Room


# Minor data cleaning
ID_13_fix <- function(ID) {
  if(ID == "13") {
    return("013")
  } else {
    return(ID)
  }
}

# Go into each data frame that is mapped into the All_Data and identify UserID that is equal to 13.
# For those files that are (Go/NoGo and N-back), change all of the rows in UserID to 013

Tibbles <- list()

for(ii in 1:nrow(All_Data)) {
  Current_UserID <- All_Data$UserID[ii]
  Current_tibble <- All_Data$Tibbles[[ii]]
  
  if(Current_UserID == "13") {
  
    Current_tibble$UserID   <- sapply(Current_tibble$UserID,ID_13_fix)
      
    Tibbles[[ii]] <- Current_tibble
    
  } else {
    
    Tibbles[[ii]] <- Current_tibble
  }
  
}



All_Data$Tibbles <- map(Tibbles,as_tibble)

# Now change it on the actual main data UserID
All_Data$UserID <- sapply(All_Data$UserID, ID_13_fix)



############ Introducing Dates #################

# Load in ONR Visitor Log Data
Visitor_Log_Full <- read_excel("~/ONR/Visitor_log/Visiting_Log_must_manually_update.xlsx")

# Create dataframe with variables that are needed
Visitor_Log <- data.frame(ID = Visitor_Log_Full$ID,
                          Day0 = Visitor_Log_Full$`Day 0 Date`,
                          Day1 = Visitor_Log_Full$`Day 1 Date`,
                          Day2 = Visitor_Log_Full$`Day 2 Date`,
                          Day3 = Visitor_Log_Full$`Day 3 Date`) %>% as_tibble()

# Change the variables into characters
Visitor_Log$Day0 <- as.character(Visitor_Log$Day0)
Visitor_Log$Day1 <- as.character(Visitor_Log$Day1)
Visitor_Log$Day2 <- as.character(Visitor_Log$Day2)
Visitor_Log$Day3 <- as.character(Visitor_Log$Day3)

# Replace all NAs to "0000-00-00" so code below can work
Visitor_Log[is.na(Visitor_Log)] <- "0000-00-00"

# Use dates from Visitor Log to determine which day we have data for in VR tasks
unique(All_Data$StartDate)

# CODE WILL BREAK IF WE HAVE DATA FOR AN ID (IN VR) THAT IS NOT PRESENT IN VISITOR LOG

Day <- list()
row_num <- list()


for(ii in 1:nrow(All_Data)) {
  Current_ID_row <- All_Data$UserID[ii]  
  Current_date_row <- All_Data$StartDate[ii]

  Visitor_Log_row <- Visitor_Log %>%
    filter(ID == Current_ID_row)
  
  row_num[ii] <- ii
  
  if(Current_date_row == Visitor_Log_row$Day0[1]) {
    Day[ii] <- "Day0"
    
  } else if(Current_date_row == Visitor_Log_row$Day1[1]) {
    Day[ii] <- "Rest"
    
  } else if(Current_date_row == Visitor_Log_row$Day2[1]) {
    Day[ii] <- "Exercise"
    
  } else if(Current_date_row == Visitor_Log_row$Day3[1]) {
    Day[ii] <- "No Sleep"
    
  } else {
    Day[ii] <- "Error"
    
  }
  
}


# Add which day belongs to which dataset
All_Data$Day <- unlist(Day)


# Add the days inside the house
Tibbles <- list()

for(ii in 1:nrow(All_Data)) {
  
  Current_Tibble <- All_Data$Tibbles[[ii]]
  Current_Tibble$Day <- rep(All_Data$Day[ii],nrow(Current_Tibble))
  Current_Tibble <- Current_Tibble %>% select(Day, everything())
  Tibbles[[ii]] <- Current_Tibble
  
}

All_Data$Tibbles_withDay <- map(Tibbles,as_tibble)


# Substr the StartDate
Tibbles <- list()

for(ii in 1:nrow(All_Data)) {
  
  Current_Tibble <- All_Data$Tibbles_withDay[[ii]]
  Current_Tibble$StartDate <- substr(rep(All_Data$StartDate[1],nrow(Current_Tibble)),1,10)
  Current_Tibble$StartDate <- as.character(Current_Tibble$StartDate)
  Current_Tibble <- Current_Tibble %>% select(Day, everything())
  Tibbles[[ii]] <- Current_Tibble
  
}

All_Data$Tibbles_withDay <- map(Tibbles,as_tibble)



# Add VR room variable to Tibbles
Tibbles <- list()

for(ii in 1:nrow(All_Data)) {
  
  Current_Tibble <- All_Data$Tibbles_withDay[[ii]]
  Current_Tibble$VR_Room  <- substr(rep(All_Data$VR_Room[ii],nrow(Current_Tibble)),1,10)
  Current_Tibble$VR_Room  <- as.character(Current_Tibble$VR_Room )
  Current_Tibble <- Current_Tibble %>% select(Day, everything())
  Tibbles[[ii]] <- Current_Tibble
  
}

All_Data$Tibbles_withDay_VR_Room <- map(Tibbles,as_tibble)



# Split the data based on their task
All_GnG <- All_Data %>%
  filter(Task_Type %in%  "Go/NoGo")

All_Nback <- All_Data %>%
  filter(Task_Type %in%  "N-back")

# Combine all the tibbles into one data frame
All_GnG_Tibbles <- do.call(rbind, as.list(All_GnG$Tibbles_withDay_VR_Room))
All_Nback_Tibbles <- do.call(rbind, as.list(All_Nback$Tibbles_withDay_VR_Room))




# Quality control
table(All_GnG_Tibbles$Day)
table(All_Nback_Tibbles$Day)

table(All_GnG_Tibbles$VR_Room)
table(All_Nback_Tibbles$VR_Room)


# IDs with errors?
All_GnG_Tibbles %>% filter(Day == "Error") %>% select(UserID) %>% unique()
All_Nback_Tibbles %>% filter(Day == "Error") %>% select(UserID) %>% unique()


# export data
# set working directory
setwd("~/ONR/GnG_and_Nback_behavioral_data_preprocessing/Processed Data")

#export the files
export(All_GnG_Tibbles, "GnG_Performance.xlsx")
export(All_Nback_Tibbles, "Nback_Performance.xlsx")