---
title: "Analyzing Math and Working Memory across Age"
author: "Leandro Ledesma"
date: "2024-03-26"
output: html_document
---

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(rempsyc) # Create APA tables with nice_table()
library(flextable, exclude = "footnote") # Adds aesthetic functions to the table from above
library(kableExtra) # Extra table formatting type
library(corrtable) # Potential correlation matrix code
library(broom) # Converts regression outputs into dataframes using the tidy() function
library(car) # To calculate VIF
library(broom)
library(ggpubr) # ggarrange
library(quantreg) # rq (quantile regression)
library(MASS, exclude = "select") # This package is loaded with QuantPsyc, must exclude "select" or you wont be able to use it. 
library(QuantPsyc) # Can use the lm.beta function to calculate the standardized betas
```


### Load in the data

```{r load in the data, warning = FALSE}
# set working directory
setwd("~/OVC/ZAT")

# load in the data
ZAT.math <- read.csv("ZAT.math.scored.csv")
ZAT.RC <- read.csv("ZAT.RC.scored.csv")

# set working directory
setwd("~/OVC/Demographics")

# load in the data
demo <- read.csv("Demographics.scored.csv")

# data cleaning
demo <- demo %>%
  select(OVC_ID = ID, 
         Sex,
         Child_age)

# set working directory
setwd("~/OVC/SES")

# load in the data
SES <- read_excel("SES.xlsx")

# data cleaning
SES <- SES %>%
  select(OVC_ID, 
         Weighted_SES_index)

# set working directory
setwd("~/OVC/Physical Data")

# load in the data
Physical <- read.csv("Physical.scored.csv")

# data cleaning
Physical <- Physical %>%
  select(OVC_ID, 
         BMI,
         Head_Circum_cm) %>%
  mutate(OVC_ID = as.numeric(OVC_ID)) %>%
  filter(BMI < 40)

# Set working directory
setwd("~/OVC/LDSpan")

# load in the data
LDSpan.num <- read.csv("LDSpan.num.scored.csv")

# data cleaning
LDSpan.num <- LDSpan.num %>%
  select(OVC_ID = ID,
         LDSpan.num.Score)

# load in the data
LDSpan <- read.csv("LDSpan.scored.csv")

# data cleaning
LDSpan <- LDSpan %>%
  select(OVC_ID = ID,
         LDSpan.scored)

```


### Combine the dataset into one


```{r combine the dataset into one}
# merged dataset
data <-
  ZAT.math %>%
  full_join(ZAT.RC, by = "OVC_ID") %>%
  full_join(demo, by = "OVC_ID") %>%
  full_join(SES, by = "OVC_ID") %>%
  full_join(Physical, by = "OVC_ID") %>%
  full_join(LDSpan.num, by = "OVC_ID") %>%
  full_join(LDSpan, by = "OVC_ID") 

dim(data)

```


### Missing data

Identifying the number of missing data per variable and then keeping only rows with no missing data. 

```{r missing data}
print("The number of missing data for each variable is:")
colSums(is.na(data))

# Remove any rows with missing data
data2 <- data %>%
  filter(complete.cases(.))

print("The number of missing data for each variable is:")
colSums(is.na(data2))

dim(data2)

```

### Descriptive statistics of our predictors and covariates

```{r descriptive statistics of our predictor and outcome variable}
# Table on continuous predictors
continuous <- data2 %>%
  select(RC.Score,
         Child_age,
         Math.Score,
         Weighted_SES_index,
         BMI,
         LDSpan.scored)

# We nee to have our continuous variables in one column
continuous.long <- continuous %>%
  stack() %>%
  select(Predictors = ind,
         values = values)

# Obtain the descriptive statistics from the long format demographic sheet
continuous.summarize <- continuous.long %>%
  group_by(Predictors) %>%
  summarize(Mean = round(mean(values),2),
            Med. = round(median(values),2),
            SD = round(sd(values),2),
            Min. = round(min(values),2),
            Max. = round(max(values),2),
            N = nrow(continuous))

# Table cleaning
continuous.summarize <- continuous.summarize %>%
  mutate(Predictors = factor(Predictors, 
                             labels = c(
                                        "Reading Comprehension (ZAT)",
                                        "Age",
                                        "Math Score (ZAT)",
                                        "Weighted SES",
                                        "BMI",
                                        "Verbal WM (Digit & Letter)")))

# Create an APA style table with the results
continuous.summarize %>%
  nice_table(title = "Descriptive Statistics of the Continuous Variables") %>%
  bold(part = "header")


# Create a table for categorical variables
sex.df <- data.frame(table(data2$Sex)) %>%
  pivot_wider(names_from = Var1, values_from = Freq) %>%
  mutate(N = `F` + M) %>%
  mutate(Percent = round(M/N,2)*100) %>%
  select(-`F`, Frequency = M, Percent, N)


# Stack the tables
categorical.df <-sex.df %>%
  mutate(`_` = "Male",
         Percent = paste(Percent,"%",sep="")) %>%
  select(`_`, Frequency, Percent, N)

# Create an APA style table with the results
categorical.df %>%
  nice_table(title = "Descriptive Statistics of Sex") %>%
  bold(part = "header")
```


## Check for assumptions

### Assumption 1: Normality (outcome variable)

```{r checking assumption 1}
# Create a histogram of the outcome variable
data2 %>%
  ggplot(aes(x = LDSpan.num.Score)) +
  geom_histogram(fill = "white",
                 color = "black",
                 bins = 12) +
  #scale_y_continuous(expand = c(0,0), limits = c(0,200)) + 
  theme_classic() +
  labs(title = "A Histogram of Digit Span",
       x = "Digit Span Score",
       y = "Frequency",
       #caption = "Figure 1: A histogram of graduation rates",
       caption = bquote(bold("Figure 1:") ~ "A histogram of Digit Span performance.")) +
  theme(plot.title = element_text(size = 14,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, 
                                  face = "bold"),
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 13,
                                    hjust = 0))

# Use a Shapiro Wilk's test on graduation rates. 
shapiro.test(data2$LDSpan.num.Score)
```

### Assumption 2: Linearity

```{r checking assumption of linearity}
# Create a scatterplot between the predictor and outcome variable
data2 %>%
  ggplot(aes(x = Math.Score, y = LDSpan.num.Score)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Scatterplot of ZAT Math Scores and Digit Span Performance",
       x = "ZAT Math Scores",
       y = "Digit Span Performance",
  caption = bquote(bold("Figure 2:") ~ "A scatterplot of math ability and digit span performance.")) +
  theme_bw() +
  theme(plot.title = element_text(size = 15,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 15,
                                    hjust = 0))

data2 %>%
  ggplot(aes(x = Child_age, y = LDSpan.num.Score)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Scatterplot of Age and Digit Span Performance",
       x = "Age (years)",
       y = "Digit Span Performance",
  caption = bquote(bold("Figure 3:") ~ "A scatterplot of age and digit span performance.")) +
  theme_bw() +
  theme(plot.title = element_text(size = 15,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 15,
                                    hjust = 0))

```

### Assumption 3: Homoscedasticity


```{r assumption 3 homoscedasticity}
# Create the boxplots from predictor quintiles
# Part 1: Obtain the quintiles of price and save it as a dataframe
quantile.df <- quantile(data2$Math.Score, probs = seq(0, 1, .2)) %>%
  data.frame(cut.off =.)

# Part 2: Create a new categorical variable defining which quintile the data belongs to
data.quant <- data2 %>%
  select(x = Math.Score,
         y = LDSpan.num.Score) %>%
  arrange(x) %>%
  mutate(quant.x = ifelse(x < quantile.df$cut.off[2], "Q1",
                          ifelse(x < quantile.df$cut.off[3], "Q2",
                                 ifelse(x < quantile.df$cut.off[4], "Q3",
                                        ifelse(x < quantile.df$cut.off[5], "Q4", "Q5")))))


# Plot the quintiles of x as boxplots + jitter for our y
data.quant %>%
  ggplot(aes(x = quant.x, y = y)) +
  geom_boxplot() +
  theme_classic() +
  labs(title = "Digit Span Scores Grouped by ZAT Math Quintiles",
       x = "ZAT Math Quintiles",
       y = "Digit Span",
       caption = bquote(bold("Figure 4:") ~ "Box plots of digit span by ZAT math quintiles.")) +
  theme(plot.title = element_text(size = 15,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 15,
                                    hjust = 0))


# Create the boxplots from predictor quintiles
# Part 1: Obtain the quintiles of price and save it as a dataframe
quantile.df <- quantile(data2$Child_age, probs = seq(0, 1, .2)) %>%
  data.frame(cut.off =.)

# Part 2: Create a new categorical variable defining which quintile the data belongs to
data.quant <- data2 %>%
  select(x = Child_age,
         y = LDSpan.num.Score) %>%
  arrange(x) %>%
  mutate(quant.x = ifelse(x < quantile.df$cut.off[2], "Q1",
                          ifelse(x < quantile.df$cut.off[3], "Q2",
                                 ifelse(x < quantile.df$cut.off[4], "Q3",
                                        ifelse(x < quantile.df$cut.off[5], "Q4", "Q5")))))


# Plot the quintiles of x as boxplots + jitter for our y
data.quant %>%
  ggplot(aes(x = quant.x, y = y)) +
  geom_boxplot() +
  theme_classic() +
  labs(title = "Digit Span Scores Grouped Age Quintiles",
       x = "Age Quintiles",
       y = "Digit Span",
       caption = bquote(bold("Figure 5:") ~ "Box plots of digit span by age quintiles.")) +
  theme(plot.title = element_text(size = 15,
                                  hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 15,
                                    hjust = 0))
```

Assumption 4: Independence


```{r checking assumption of independence}
covariates <- data2 %>%
  select(-c(OVC_ID, 
            Head_Circum_cm, 
            LDSpan.num.Score, 
            LDSpan.scored, 
            Math.Score, 
            Sex,
            Child_age)) 

# dimensions of covariates
dim(covariates)

# Remove any missing data in our covariates
covariates <- drop_na(covariates)

# dimensions of covariates (Dropped 15 rows)
dim(covariates)

# Run a correlation matrix, but it is difficult to see- so let's change it a bit
# 1. convert the correlation matrix into a data.frame
cor.df <- cor(covariates) %>%
  data.frame() 

# 2. convert it to long then return only the correlations larger than .6 (Not including 1.00)
cor.df %>%
  mutate(variables = row.names(cor.df)) %>%
  pivot_longer(-variables) %>%
  filter(value > .6 & value != 1) %>%
  arrange(value) %>%
  mutate(value = round(value,2)) %>%
  kbl() %>%
  kable_paper(bootstrap_options = "striped", full_width = F)


```


### Creating a correlation Matrix of our Covariates

```{r creating a correlation matrix of our covariates}
# Change the name of the covariates so it's prettier
names(covariates) <- c("ZAT Reading Comprehension",
                       "Weighted SES",
                       "BMI")

# Create a correlation matrix
correlation.matrix.df <- correlation_matrix(covariates) %>%
  data.frame()

# Reintroduce names to the correlation matrix dataframe
names(correlation.matrix.df) <- names(covariates)

# Print our APA correlation matrix
correlation.matrix.df %>%
  kbl(caption = "Table 1: Correlation matrix of our continuous covariates") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  footnote(general = c("*** indicates p <.001", "** indicates p <.01 ","* indicates p <.05 "))


```


### Which variables correlate with our outcome variable?


```{r correlation matrix including outcome and predictors}
# This time we will do it for all 
exhaustive.continuous.predictors <- data2 %>%
  select(-c(OVC_ID,
            Sex, 
            Head_Circum_cm, 
            LDSpan.scored)) 

# dimensions of covariates
dim(exhaustive.continuous.predictors)

# Remove any missing data in our covariates
exhaustive.continuous.predictors <- drop_na(exhaustive.continuous.predictors)

dim(exhaustive.continuous.predictors)

# Data cleaning, have digit span and ZAT math and age as the first three variables in the dataframe
exhaustive.continuous.predictors <- exhaustive.continuous.predictors %>%
  select(LDSpan.num.Score, Math.Score, Child_age, everything())

# Run a correlation matrix, but it is difficult to see- so let's change it a bit
# 1. convert the correlation matrix into a data.frame
cor.df <- cor(exhaustive.continuous.predictors) %>%
  data.frame() 

# 2. convert it to long then return only the correlations larger than .6 (Not including 1.00)
cor.df %>%
  mutate(variables = row.names(cor.df)) %>%
  pivot_longer(-variables) %>%
  filter(value > .6 & value != 1) %>%
  arrange(value) %>%
  mutate(value = round(value,2)) %>%
  kbl() %>%
  kable_paper(bootstrap_options = "striped", full_width = F)

```

### Creating a correlation Matrix - Part 2

```{r creating a correlation matrix part 2}
# Create a correlation matrix
correlation.matrix.df <- correlation_matrix(exhaustive.continuous.predictors) %>%
  data.frame()

# Make the names prettier
names(correlation.matrix.df) <- c("Digit Span",
                                  "ZAT Math",
                                  "Age",
                                  "Reading Comprehension",
                                  "Weighted SES",
                                  "BMI")

# Need to rearrange the order or rows to show strength descending order
# Will have to get creative here
correlation.matrix.df.orderd <- correlation.matrix.df %>%
  mutate(order = as.numeric(gsub("\\D", "", `Digit Span`))) %>%
  arrange(desc(order)) %>%
  select(-order)

# Change the row names to match the variable names
row.names(correlation.matrix.df) <- names(correlation.matrix.df.orderd)

# Print our APA correlation matrix
correlation.matrix.df %>%
  kbl(caption = "Table 2: Correlation matrix of continuous predictors, covariates, and the outcome") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  footnote(general = c("*** indicates p <.001", "** indicates p <.01 ","* indicates p <.05 "))

```


### Part 1: Run a multiple regression to investigate math ability and digit span performance

Based on the results from above, we have decided to keep the following predictors/covariates

- Zat math
- Age
- Weighted SES
- Child Sex
- Outcome: Digit Span

```{r runing a multiple regression to investigate math ability}
# Change scipen option back to normal
options(scipen = 0)

# Change our categorical variable into a dummy variable
data2$SexM <- ifelse(data2$Sex == "M", 1, 0)

# What is the outcome variable?
Outcome = "digit span"

# Create the model
model1 <- lm(LDSpan.num.Score ~ Math.Score + Child_age + Math.Score*Child_age + Weighted_SES_index + SexM , data2)

#################################### Part 1: Obtaining the R^2 and more #############################

# Save the model summary into an object
model.summary <- summary(model1)

# Creating the foot note for the model
# Create line 1
line.1 <- paste("Residual standard error:",round(model.summary$sigma,4),", on",model1$df.residual,"degrees of freedom")

# Create line 2
line.2 <- paste("Multiple R-squared:",round(model.summary$r.squared,4),", Adjusted R-squared: ",round(model.summary$adj.r.squared,5), sep =" ")

# Calculate p value for line 3
regression.p.value <- pf(model.summary$fstatistic[1], model.summary$fstatistic[2], model.summary$fstatistic[3], lower.tail = FALSE)

# Create line 3
line.3 <- paste("F-statistic: ",round(model.summary$fstatistic[1],2)," on ",round(model.summary$fstatistic[2],4)," and ",round(model.summary$fstatistic[3],4)," DF,  p-value:", regression.p.value)


################################ Part 2: Introduce VIF and more ######################3


# Save the models as a dataframe
model1.df <- broom::tidy(model1)

# Calculate the standardized betas from the model and save them into the model.df
model1.df$standardized.beta <- c(NA,lm.beta(model1))

# Calculate the tolerance
model1.df$Tolerance <- c(NA,1/vif(model1))

# Calculate the VIF of the model
model1.df$VIF <- c(NA,vif(model1))

```

### Make the multiple regression look prettier

```{r make the regression using reading comprehension look pretty}
# Round only numeric variables in the dataframe
model1.df.rounded <- model1.df %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))

# Add a function to clean up the p-values
model1.df.rounded <- model1.df.rounded %>%
  mutate(p.value = case_when(p.value < .001 ~ paste(format(p.value, nsmall = 2),"***",sep=""),
                              p.value < .01 ~ paste(format(p.value, nsmall = 2),"**",sep=""),
                              p.value < .05 ~ paste(format(p.value,nsmall = 2),"*",sep=""),
                              TRUE ~ paste(p.value)))

# Change the term values
model1.df.rounded$term <- c("(Intercept)",
                            "ZAT Math",
                            "Age",
                            "Weighted SES",
                            "SexM",
                            "ZAT Math x Age")

# Change the names of the model output
names(model1.df.rounded) <- c("Model", 
                              "Beta",
                              "Std. Error",
                              "t",
                              "Sig.",
                              "Std Beta",
                              "Tolerance",
                              "VIF")

# Print to check the cleaning
model1.df.rounded %>%
  kbl(caption = paste("Table 3: Initial multiple regression predicting",Outcome),
      align = "lccclcccc",
      font = 20) %>%
  kable_classic_2(font_size = 16,
                  full_width = FALSE,
                  html_font = "Times New Roman") %>%
  gsub("font-size: initial !important;", "font-size: 14pt !important;", .) %>% # Title font size!
  row_spec(0,bold=TRUE) %>%
  footnote(general_title = "*** indicates p <.001; ** indicates p <.01; * indicates p <.05",
           general =  c(line.1, 
                        line.2, 
                        line.3))

```


### Part 2: Run a multiple regression to investigate math ability and digit span performance (remove not significant predictors)

```{r runing a multiple regression to investigate math ability part 2}


# Create the model
model2 <- lm(LDSpan.num.Score ~ Math.Score + Child_age + Math.Score*Child_age  , data2)

#################################### Part 1: Obtaining the R^2 and more #############################

# Save the model summary into an object
model.summary <- summary(model2)

# Creating the foot note for the model
# Create line 1
line.1 <- paste("Residual standard error:",round(model.summary$sigma,4),", on",model2$df.residual,"degrees of freedom")

# Create line 2
line.2 <- paste("Multiple R-squared:",round(model.summary$r.squared,4),", Adjusted R-squared: ",round(model.summary$adj.r.squared,5), sep =" ")

# Calculate p value for line 3
regression.p.value <- pf(model.summary$fstatistic[1], model.summary$fstatistic[2], model.summary$fstatistic[3], lower.tail = FALSE)

# Create line 3
line.3 <- paste("F-statistic: ",round(model.summary$fstatistic[1],2)," on ",round(model.summary$fstatistic[2],4)," and ",round(model.summary$fstatistic[3],4)," DF,  p-value:", regression.p.value)


################################ Part 2: Introduce VIF and more ######################3


# Save the models as a dataframe
model2.df <- broom::tidy(model2)

# Calculate the standardized betas from the model and save them into the model.df
model2.df$standardized.beta <- c(NA,lm.beta(model2))

# Calculate the tolerance
model2.df$Tolerance <- c(NA,1/vif(model2))

# Calculate the VIF of the model
model2.df$VIF <- c(NA,vif(model2))

```


### Make the multiple regression look prettier (part 2)

```{r make the regression using reading comprehension look pretty part 2}
# Round only numeric variables in the dataframe
model2.df.rounded <- model2.df %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))

# Add a function to clean up the p-values
model2.df.rounded <- model2.df.rounded %>%
  mutate(p.value = case_when(p.value < .001 ~ paste(format(p.value, nsmall = 2),"***",sep=""),
                              p.value < .01 ~ paste(format(p.value, nsmall = 2),"**",sep=""),
                              p.value < .05 ~ paste(format(p.value,nsmall = 2),"*",sep=""),
                              TRUE ~ paste(p.value)))

# Change the term values
model2.df.rounded$term <- c("(Intercept)",
                            "ZAT Math",
                            "Age",
                            "ZAT Math x Age")

# Change the names of the model output
names(model2.df.rounded) <- c("Model", 
                              "Beta",
                              "Std. Error",
                              "t",
                              "Sig.",
                              "Std Beta",
                              "Tolerance",
                              "VIF")

# Print to check the cleaning
model2.df.rounded %>%
  kbl(caption = paste("Table 4: Final multiple regression predicting",Outcome),
      align = "lccclcccc",
      font = 20) %>%
  kable_classic_2(font_size = 16,
                  full_width = FALSE,
                  html_font = "Times New Roman") %>%
  gsub("font-size: initial !important;", "font-size: 14pt !important;", .) %>% # Title font size!
  row_spec(0,bold=TRUE) %>%
  footnote(general_title = "*** indicates p <.001; ** indicates p <.01; * indicates p <.05",
           general =  c(line.1, 
                        line.2, 
                        line.3))

```


### Part 3: Run a multiple regression that is CENTERED to investigate math ability and digit span performance (remove not significant predictors)

The goal here is to remove multicollinearity AND to make better sense of the y-intercept

```{r runing a multiple regression to investigate math ability part 3}
# Add the centered version of our predictors.
data2$Math.Score.c <- data2$Math.Score - mean(data2$Math.Score)
data2$Child_age.c <- data2$Child_age - mean(data2$Child_age)

# Create the model
model3 <- lm(LDSpan.num.Score ~ Math.Score.c * Child_age.c  , data2)

#################################### Part 1: Obtaining the R^2 and more #############################

# Save the model summary into an object
model.summary <- summary(model3)

# Creating the foot note for the model
# Create line 1
line.1 <- paste("Residual standard error:",round(model.summary$sigma,4),", on",model3$df.residual,"degrees of freedom")

# Create line 2
line.2 <- paste("Multiple R-squared:",round(model.summary$r.squared,4),", Adjusted R-squared: ",round(model.summary$adj.r.squared,5), sep =" ")

# Calculate p value for line 3
regression.p.value <- pf(model.summary$fstatistic[1], model.summary$fstatistic[2], model.summary$fstatistic[3], lower.tail = FALSE)

# Create line 3
line.3 <- paste("F-statistic: ",round(model.summary$fstatistic[1],2)," on ",round(model.summary$fstatistic[2],4)," and ",round(model.summary$fstatistic[3],4)," DF,  p-value:", regression.p.value)


################################ Part 2: Introduce VIF and more ######################3


# Save the models as a dataframe
model3.df <- broom::tidy(model3)

# Calculate the standardized betas from the model and save them into the model.df
model3.df$standardized.beta <- c(NA,lm.beta(model3))

# Calculate the tolerance
model3.df$Tolerance <- c(NA,1/vif(model3))

# Calculate the VIF of the model
model3.df$VIF <- c(NA,vif(model3))

```


### Make the multiple regression look prettier (part 2)

```{r make the regression using reading comprehension look pretty part 3}
# Round only numeric variables in the dataframe
model3.df.rounded <- model3.df %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))

# Add a function to clean up the p-values
model3.df.rounded <- model3.df.rounded %>%
  mutate(p.value = case_when(p.value < .001 ~ paste(format(p.value, nsmall = 2),"***",sep=""),
                              p.value < .01 ~ paste(format(p.value, nsmall = 2),"**",sep=""),
                              p.value < .05 ~ paste(format(p.value,nsmall = 2),"*",sep=""),
                              TRUE ~ paste(p.value)))

# Change the term values
model3.df.rounded$term <- c("(Intercept)",
                            "ZAT Math",
                            "Age",
                            "ZAT Math x Age")

# Change the names of the model output
names(model3.df.rounded) <- c("Model", 
                              "Beta",
                              "Std. Error",
                              "t",
                              "Sig.",
                              "Std Beta",
                              "Tolerance",
                              "VIF")

# Print to check the cleaning
model3.df.rounded %>%
  kbl(caption = paste("Table 4: Final multiple regression predicting",Outcome),
      align = "lccclcccc",
      font = 20) %>%
  kable_classic_2(font_size = 16,
                  full_width = FALSE,
                  html_font = "Times New Roman") %>%
  gsub("font-size: initial !important;", "font-size: 14pt !important;", .) %>% # Title font size!
  row_spec(0,bold=TRUE) %>%
  footnote(general_title = "*** indicates p <.001; ** indicates p <.01; * indicates p <.05",
           general =  c(line.1, 
                        line.2, 
                        line.3))

```



### Visualize the interaction

```{r visualize the interaction for reading comprehension}
library(interactions)

interact_plot(model = model2, pred = Math.Score, modx = Child_age)

library(psych)
describe(data2$Child_age)
```

### Another approach to visualizing the interaction

```{r visualing interaction part 2}
library(rockchalk)

plotPlane(model2, 'Child_age', 'Math.Score',
          phi = 10, theta = -55, ticktype = "detailed")

```



### Visualizing residuals

```{r visualizing residuals for reading comprehension}
# introduce the residuals into the datasets
data2$studentized.residuals <- rstudent(model2)

# Create a histogram of the outcome variable
residual.plot1 <- data2 %>%
  ggplot(aes(x = studentized.residuals)) +
  geom_histogram(fill = "white",
                 color = "black",
                 bins = 15) +
  scale_y_continuous(expand = c(0,0), limits = c(0,250)) + 
  theme_classic() +
  labs(title = "A histogram of Studentized Residuals",
       x = "Studentized Residuals",
       y = "Frequency",
       caption = bquote(bold("Figure 6:") ~ "A histogram of Studentized Residuals.")) +
  theme(plot.title = element_text(size = 11,
                                  hjust = 0.5),
        axis.title = element_text(size = 10, face = "bold"),
        axis.text = element_text(size = 10),
        plot.caption = element_text(size = 8,
                                    hjust = 0))

# Create a QQ plot
residual.plot2 <- data2 %>%
  ggplot(aes(sample= studentized.residuals)) +
  geom_qq() +
  stat_qq_line() +
  theme_classic() +
  labs(title = "QQ Plot",
       x = "Theoretical Quantiles",
       y = "Frequency",
       caption = bquote(bold("Figure 7:") ~ "A QQ Plot of studentized residuals and theoretical quantiles.")) +
  theme(plot.title = element_text(size = 11,
                                  hjust = 0.5),
        axis.title = element_text(size = 10, face = "bold"),
        axis.text = element_text(size = 10),
        plot.caption = element_text(size = 8,
                                    hjust = 0))

# Plot the graphs
ggarrange(residual.plot1, residual.plot2)


```

### Post-estimation techniques (homoscedasticity)

```{r homoscedasiticyt, out.width= "70%"}
library(lmtest) # bptest() function for homoscedasticity

# Run the Breusch-Pagan test
bptest(model2)

# Add fitted values and standardized residuals
data2$fitted.values <- model2$fitted.values
data2$standardized.residuals <- rstandard(model2)

# Plot to see if there is a problem with homoscedasticity
data2 %>%
  ggplot(aes(x = fitted.values, y=standardized.residuals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  labs(title = "Standardized Residuals vs Fitted Values ",
       x = "Fitted Values",
       y = "Standardized Residuals",
       caption = bquote(bold("Figure 8:") ~ "Investigating homoscedasticity")) +
  theme(plot.title = element_text(size = 18,
                                  hjust = .5),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14),
        plot.caption = element_text(size = 15,
                                    hjust = 0))
```



### Create a correlation matrix of predictors

```{r lastly create a correlation matrix}
data3 <- data2 %>%
  mutate(Math.Age.Interaction = Math.Score * Child_age) %>%
  select(Math.Score, Child_age, Math.Age.Interaction, LDSpan.num.Score)

library(GGally)

data3 %>%
  rename(`ZAT Math` = Math.Score, `Age` = Child_age, `ZAT Math x Age` = Math.Age.Interaction, `Digit Span` = LDSpan.num.Score) %>%
  ggpairs(title = "Scatter Plot Matrix of Predictors in Final Model") +
  labs(caption = bquote(bold("Figure 9:") ~ "Scatter plot matrix of predictors in the final model")) +
  theme(plot.title = element_text(size = 18,
                                  hjust = .5),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14),
        plot.caption = element_text(size = 15,
                                    hjust = 0))
````