---
title: "Analyzing effects of Color on Performance in the Color Saber Match Condition of the VR Go/No-Go"
author: "Leandro Ledesma"
date: "2024-02-21"
output: html_document
---


### Overview 

This is a research idea by Candace Martinez that investigates if the wavelength properties of color  affect performance in the Color Saber Match Condition of the Go/No-Go (GnG). In a quick recap, color is perceived from wavelengths of light, and several studies showed that there is an association between color wavelength and performance in different types of cognitive related tasks. For our purposes, we have a version of the GnG called the Color Saber Matching condition. In this version, we have 210 trials of which 168 trials (80%) are Go trials and 42 trials (20%) are No-Go trials. The Go trials are defined as conditions where the presented stimulus, which can be one of the following colors: purple, blue, green, yellow, red, matches one of the colors of the sabers. In contrast, the No-Go trials is a condition where the color of the stimulus does not match the color of the sabers. For further context, this GnG task is being implemented in a virtual reality (VR) environment. Thus, the participant will be wielding a joy stick in both hands that function as virtual sticks (sabers) to hit stimuli during the Go-trials.

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
require(tidyverse)
library(readxl)

library(lme4)
library(lmerTest) 
library(lsr) # Shows the p values when you run the ANOVAS
library(emmeans) # Lets you do follow up tests on the ANOVAs
library(purrr)
library(simr) # Post hoc tests on our linear mixed effects model
library(ggplot2)

```

### Load in the data

The following dataset is somewhat independent of the other ONR data I load into my analysis. This is semi-processed data that needs a special script to be run to get the same results as when it was first run sometime late 2023. This is data that was analyzed and written for a manuscript that then needed a couple of revisions. Thus, we need to keep the same sample size to reduce the number of changes needed for revision.


```{r load in the data, warning = FALSE}
# Set working directory
setwd("C:/Users/lledesma.TIMES/Documents/ONR/GnG_and_Nback_behavioral_data_preprocessing/Processed Data")

GnG <- read_excel("GnG_Performance.xlsx")
```

### Data Cleaning

These data were already somewhat cleaned in a previous (no longer existing) script. Thus we will need to do a little extra cleaning to remove practice days and any days labeled incorrectly ('error') from the visitor log.

Lastly, we only want to keep only data from VR Room 1. 

```{r data cleaning}
GnG_VR_Room_one <- GnG %>%
  filter(VR_Room == "ONR")

# GnG remove Day 0s
GnG_removed_Day0 <- GnG_VR_Room_one %>%
  filter(Day != "Day0")

# GnG remove Errors (Don't question this)
GnG_removed_error <- GnG_removed_Day0 %>%
  filter(Day != "Error")

Saber_Color <- GnG_removed_error %>%
  filter(Day == "Rest" & BlockName =='Saber Color Matching')

```


### Creating some new variables

We are now going to create some variables (I don't remember if they are any useful but we are not deviating from the original code).

```{r creating some new variables}
# Create a function that identifies the two different types of errors ONLY in GO Trials
# Go Trial Error 1 = Miss
# Go Trial Error 2 = Poor Arrow Accuracy

Behavior_Type_fun <- function(Trial_Type,Correct,Reacted) {
  if(Trial_Type == "G12") {
    
   if(Correct == "1" & Reacted == "1") {
     return("Hit")
     
   } else if (Correct == "0" & Reacted == "1") {
    return("Poor Accuracy")
     
   } else if (Correct == "0" & Reacted == "0") {
    return("Miss")
     
   }
    
  } else if(Trial_Type == "G11") {
    
    if(Correct == "1" & Reacted == "0") {
      return("Inhibition")
      
    } else if (Correct == "0" & Reacted == "1") {
      return("Failed Inhibition")
    
  }
  
}

}

# Run the function through a for loop
Behavior_Type <- list()

for(ii in 1:nrow(Saber_Color)) {
  
  Trial_Type <- Saber_Color$ItemID[[ii]]
  Correct <- Saber_Color$Correct[[ii]]
  Reacted <- Saber_Color$Reacted[[ii]]
  
  Behavior_Type[[ii]] <- Behavior_Type_fun(Trial_Type,Correct,Reacted)
  
}

Saber_Color$Behavior_Type <- unlist(Behavior_Type)

# Create a function that makes a numeric variable of our color

color_fun <- function(x) {
  if(x == "purple") {
    return(407)
  } else if (x == "blue") {
    return(478)
  } else if (x =="green") {
    return(507)
  } else if (x =="yellow") {
    return(612)
  } else if (x =="red") {
    return(679)
  }
  
}

Saber_Color$Color_num<- sapply(Saber_Color$Color,color_fun )

```


### Extra data cleaning

```{r extra data cleaning}
# Keep only the variables we are interested in 
Final_saber_data <- Saber_Color %>%
  select(UserID,ItemID, Color,Color_num, Correct, Reacted, TimeToImpact, Behavior_Type)

# Change the values of the ItemID so it is easier to understand
Final_saber_data <- Final_saber_data %>%
  mutate(ItemID = ifelse(ItemID == "G11","No-Go","Go"))


```


### Loading demographics and IQ data

```{r adding demographincs, warning= FALSE}
# set the working directory for the demographic data
setwd("Z:/Leo/Demographics")

# import demographic data
demo <- read.csv("complete.demographics.csv")

# select the variables of interest
demo_final <- demo %>%
  select(UserID = ID, AgeYears = Age,Gender = Sex)

demo_final$AgeYears <- as.numeric(demo_final$AgeYears)

# set the working directory for the IQ data
setwd("Z:/Leo/CFIT")

# Import IQ data
IQ_data <- read.csv("CFIT.raw.scores.csv")

# Clean IQ data
IQ_data <- IQ_data %>%
  select(UserID = ID, Raw_Score = CFIT)
```


### Merge the datasets into one


```{r adding IQ data}
# Merge the data into one
Final_data <- Final_saber_data %>%
  mutate(UserID = as.numeric(UserID)) %>%
  left_join(demo_final, by = "UserID") %>%
  left_join(IQ_data, by = "UserID")

```


### Reducing data dimensionality

At this point we have 15960 rows, with 210 trials for each participant that show if a trial was correct by displaying a 1 or incorrect by displaying a 0. We are now going to obtain the proportion average for each person for all of their Go and No-Go trials. Thus, each subject should have 2 (trial type; Go vs NoGo) x color (5 wavelength colors) = 10 rows. 

Additionally, we will also do the same for a separate dataframe that represents reaction time. Except, this dataset will only contain Go trials that are correct (there is no reaction time for correct No-Go trials since they are not supposed to be hit). Thus, this dataset will have 5 rows (5 wavelength colors) per subject. 

```{r reducing data dimensionality}
# Current data dimension
dim(Final_data)

# Creating a dataset for accuracy
Analysis_data2 <- Final_data %>%
  group_by(UserID, ItemID, Color_num,Raw_Score,Gender,AgeYears) %>%
  summarize(Performance = mean(Correct)) %>%
  ungroup()

# The dimensions of the accuracy dataset
dim(Analysis_data2)

# Creating a dataset for reaction time
Analysis_RT_data2 <- Final_data %>%
  filter(Correct == 1) %>%
  group_by(UserID,Color_num,Raw_Score,Gender,AgeYears) %>%
  summarize(mean_RT = mean(TimeToImpact), median_RT = median(TimeToImpact)) %>%
  ungroup()

# The dimensions of the reaction time dataset
dim(Analysis_RT_data2)

```
### Removing outliers

```{r removing potential outliers}
###### Identifying outliers for Go signals and No-Go signals
####

overall_performance_df <- Analysis_data2 %>%
  filter(ItemID == "Go") %>%
  group_by(UserID) %>%
  summarize(Overall_Performance = mean(Performance))

overall_performance_df$Overall_Performance %>% summary()
  
sd_overall_performance <- overall_performance_df$Overall_Performance %>% sd()
mean_overall_performance <-  overall_performance_df$Overall_Performance %>% mean()

outlier_IDs_Go_signals <- overall_performance_df %>%
  filter(Overall_Performance < mean_overall_performance - 2*sd_overall_performance) %>%
  select(UserID) %>%
  unlist() %>% unname()


#####
############# Remove outlier IDs
#####
`%nin%` = Negate(`%in%`)

Analysis_data2 <- Analysis_data2 %>%
  filter(UserID %nin% outlier_IDs_Go_signals)

Analysis_RT_data2 <- Analysis_RT_data2 %>%
  filter(UserID %nin% outlier_IDs_Go_signals)
```

### Quality control


```{r quality control aka inspection}
# Gender numbers
Analysis_data2 %>%
  select(UserID, Gender) %>%
  unique() %>%
  count(Gender)

# Number of participants
length(unique(Analysis_data2$UserID))

# Other stuff
Analysis_data2$AgeYears %>% mean() %>% round(2) # mean age= 22.38
Analysis_data2$AgeYears %>% sd() %>% round(2) # sd age = 3.52
Analysis_data2$Raw_Score %>% mean() %>% round(2) # mean raw IQ = 34.99
Analysis_data2$Raw_Score %>% sd() %>% round(2) # sd raw IQ = 4.9
Analysis_data2$Raw_Score %>% summary()
```

### Analyzing the data

```{r analyzing the data}
# Split data into Go and No-Go datasets
Go_trials <- Analysis_data2 %>%
  filter(ItemID == "Go")

No_Go_trials <- Analysis_data2 %>%
  filter(ItemID == "No-Go") 


lmer(Performance~ Color_num + Raw_Score + Gender + AgeYears + (1|UserID), Go_trials) %>%
    anova() # Only color is significant; Gender and Age had no effect

Go_accuracy_model <- lmer(Performance~ Color_num + Raw_Score + Gender + AgeYears + (1|UserID), Go_trials)

lmer(Performance~Color_num*Gender + Raw_Score +  + AgeYears + (1|UserID), Go_trials) %>%
  anova() # No gender interaction effect with color wavelength

Go_accuracy_model_interaction <- lmer(Performance~Color_num*Gender + Raw_Score +  + AgeYears + (1|UserID), Go_trials)

lmer(Performance~ Color_num + Raw_Score + Gender + AgeYears + (1|UserID), No_Go_trials) %>%
  anova() # Only color is significant; Gender and Age had no effect

No_Go_accuracy_model <- lmer(Performance~ Color_num + Raw_Score + Gender + AgeYears + (1|UserID), No_Go_trials)

lmer(Performance~Color_num*Gender + Raw_Score +  + AgeYears + (1|UserID), No_Go_trials) %>%
  anova() # No gender interaction effect with color wavelength

No_Go_accuracy_model_interaction <- lmer(Performance~Color_num*Gender + Raw_Score +  + AgeYears + (1|UserID), No_Go_trials) 


```


### Assumptions


```{r showing our assumptions}





````