---
title: "GnG EEG N2 Analysis"
author: "Leandro Ledesma"
date: "2024-02-29"
output: html_document
---

### Our measure of interest





### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```

### Load in the data manipulation packages first

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readxl)
library(lmtest) # bptest() function for homoscedasticity
library(kableExtra)

```

### Load in the data


```{r load in the data, warning = FALSE}
# Set working directory
setwd("M:/EEG_XDFs_Copy_2/6_GnG_EEG ERPs means")

# obtain the list names of all the files
all.file.names <- list.files()

# keep only Day 1 files
Day1.N2.files <- all.file.names[grepl(pattern = "Day1",
                       x = all.file.names)]

# Load in the files
N2.list <- list()

for(ii in 1:length(Day1.N2.files)) {

  # Load in the data
  current.data <- read.csv(Day1.N2.files[ii],
                           sep = "\t")
  
  
  # Save this transformed dataset into the list
  N2.list[[ii]] <- current.data %>% select(ERPset, everything())
  
}

# Merge into one dataset
N2.data <- N2.list %>%
  do.call(rbind,.)

```


### Data Cleaning

We have variable names that indicate 1) whether the trial was a Go (bin1) or a NoGo (bin2) trial and 2) what electrode the brain activity was recorded from. What we need to do is to extract the bin information from the variable names and have that as its own variable. 

```{r data cleaning}
# Create a dataset for Go trial activity
N2.Go.bin <- N2.data %>%
  select(ERPset, starts_with("bin1")) %>%
  pivot_longer(-ERPset, names_to = "Bin.Type", values_to = "Value")

N2.NoGo.bin <- N2.data %>%
  select(ERPset, starts_with("bin2")) %>%
  pivot_longer(-ERPset, names_to = "Bin.Type", values_to = "Value")


# Merge the datasets together
N2.long <- N2.data %>%
  pivot_longer(-ERPset, names_to = "Bins", values_to = "Value") %>%
  mutate(Trial.Type = ifelse(grepl("bin1",Bins),"Go","NoGo"),
         Channel = sub("^[^_]*_", "", Bins)) %>%
  select(ERPset, Trial.Type, Channel, Value)

```


# Graphing the data (Line graph)


```{r graphing time}
# Average all of the data by channels and Trial.Type

averaged.N2 <- N2.long %>%
  group_by(Trial.Type, Channel) %>%
  summarize(mean.Value = mean(Value))


# Graph it
averaged.N2 %>%
  ggplot(aes(x = Channel, y = mean.Value, color = Trial.Type, group = Trial.Type)) +
  geom_point() +
  stat_summary(fun = "mean",
               geom = "line") +
  theme_bw()

```

# Graphing the data part (Graphing by topographical location)

```{r graphing time2}

# Create a grouping based on topographical location

Topographical_group_fun <- function(channel) {
  case_when(
    channel %in% c("AF3", "AF4", "F3", "F4","F7", "F8", "Fz") ~ "Frontal",
    channel %in% c("FT10", "FT9","T7","T8") ~ "Temporal",
    channel %in% c("FC1", "FC2", "FC5", "FC6", "C3", "C4", "CP1", "CP2", "CP5", "CP6","CZ") ~ "Central",
    channel %in% c("P3","P4","P7","P8","Pz") ~ "Parietal",
    channel %in% c("O1","O2", "Oz") ~ "Occipital",
    TRUE ~ as.character(channel)
  )
}

N2.long <- N2.long %>%
  mutate(Topography = Topographical_group_fun(Channel),
         Topography = factor(Topography,
                             levels = c("Frontal",
                                        "Central",
                                        "Temporal",
                                        "Parietal",
                                        "Occipital")))


# Graph as bar graphs
N2.long %>%
  ggplot(aes(x = Topography, y = Value, fill = Trial.Type)) +
  stat_summary(fun = "mean",
               geom = "bar",
               position = "dodge") +
  stat_summary(fun.data = mean_cl_normal,
               geom = "errorbar",
               width = .2,
               position = position_dodge(width = .9)) +
  geom_jitter(position = position_dodge(width = .7)) +
  labs(title = "N2 between averaged Go and NoGo trials across topographical locations") +
  theme_classic()


# Graph as bar graphs
N2.long %>%
  ggplot(aes(x = Topography, y = Value, fill = Trial.Type)) +
  stat_summary(fun = "mean",
               geom = "bar",
               position = "dodge") +
  stat_summary(fun.data = mean_cl_normal,
               geom = "errorbar",
               width = .2,
               position = position_dodge(width = .9)) +
  labs(title = "N2 between averaged Go and NoGo trials across topographical locations") +
  theme_classic()

```
