---
title: "Identifying Pilot"
author: "Leandro Ledesma"
date: "2024-04-19"
output: html_document
---

---
title: "Scoring Qualtrics"
output: html_notebook
---


This contains all of the information relevant to scoring Qualtrics (with the exception of BRIEF and ASVAB).

### Universal block code settings

```{r universal block code setting}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
```

### Load in the data manipulation packages first

```{r load in packages, message= FALSE, warning = FALSE}
# Load in packages
library(tidyverse)
```


### Load in the data

```{r load in the data, echo= F, warning = FALSE}
#set working directory
setwd("C:/Users/lledesma.TIMES/Documents/ONR/Qualtrics")

# Load in the ASVAB data 
all.files <- list.files()
Qualtrics.file <- all.files[grepl("Questionnaires", all.files)]
Qualtrics <- read.csv(Qualtrics.file)

```
### Cleaning the data

Remove the first 18 rows- they are meaningless.

```{r asds}
# Start on row 18 since that is real data. 
Qualtrics2 <- Qualtrics[18:nrow(Qualtrics),]

# Remove scientific notation
options(scipen=999)

# Transform the ID variable to numeric
Qualtrics2$ID <- as.numeric(gsub("\\D", "", Qualtrics2$ID))

# Check for duplicate ID's
duplicate.IDs <- Qualtrics2 %>%
  select(ID) %>%
  count(ID) %>%
  filter(n > 1) %>%
  select(ID)

duplicate.IDs

# Remove duplicate IDs from the dataset
paste("The following IDs were removed from further analysis for having duplicates", unlist(duplicate.IDs))

# Remove the ID's that have duplicates
Qualtrics2 <- Qualtrics2 %>%
  filter(!(ID %in% unlist(duplicate.IDs)))

```


### Label members in our sample as ROTC/Military or not

```{r label them bois}
Qualtrics2 <- Qualtrics2 %>%
  mutate(Military.Background = case_when(
    
    ROTC.Exp1 == "Yes" ~ "Yes",
    ME1 == "Yes" ~ "Yes",
    TRUE ~ "No"
    
  ))

# Descriptives of our subjects with a military background
table(Qualtrics2$Military.Background)

# Filter our subjects with military experience
ROTC.Mil.exp <- Qualtrics2 %>%
  filter(Military.Background == "Yes")

# Select variables of interest
ROTC.Mil.exp.questions <- ROTC.Mil.exp %>%
  select(ID, ROTC.Exp1, ROTC.Exp2, ROTC.Exp2_4_TEXT, ROTC.Exp3, ROTC.Exp4, ROTC.Exp5, ROTC.Exp6, ROTC.Exp6_4_TEXT, ROTC.Exp7,
         ME1, ME2, ME2_6_TEXT, ME3, ME4, ME5, ME6, ME7, ME8, ME9 )
```


### Check from this sample who is still enlisted (ROTC/Military)

```{r ROTC and military groups}
### Still enlisting in either ROTC or Military
ROTC.Mil.exp.questions <- ROTC.Mil.exp.questions %>%
  mutate(Still.Enlisting = case_when(
    
    ROTC.Exp3 == "Yes" ~ "Yes",
    ROTC.Exp3 == "No, I commissioned out of the program" ~ "Yes",
    ME3 == "Yes" ~ "Yes",
    TRUE ~ "No"
  ))

# Set the working directory to include demographics
setwd("~/ONR/Demographics")

# Load in demographics data
demo <- read.csv("complete.demographics.csv")
demo.cleaned <- demo %>%
  select(ID, Sex, Age)

# Include the demographics to our military sample
ROTC.Mil.exp.questions <- ROTC.Mil.exp.questions %>%
  left_join(demo.cleaned, by = "ID")

ROTC.Mil.exp.questions %>%
  group_by(Still.Enlisting) %>%
  summarize(n = length(Still.Enlisting),
            age = mean(Age))


table(ROTC.Mil.exp.questions$Still.Enlisting)
```
### Set working directory to save

```{r save information}
# Set the directory to save this information
setwd("C:/Users/lledesma.TIMES/Documents/ONR/Pilot Sample")

# Save the dataset
write_csv(ROTC.Mil.exp.questions, "Potential.Pilot.csv")

```

