---
title: "Analyzing the PSQ"
author: "Leandro Ledesma"
date: "2024-05-12"
output: html_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)

```


```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
library(openxlsx) # To save excel files with multiple tabs
library(ggpubr)
```


```{r load in the PSQ data, warning = FALSE}
# Set working directory
setwd("~/KBB_new_2/2_behavioral_assessments/Adults/Perceived_Stigma_Questionnaire_(PSQ)/raw_data")

# Load in the data
PSQ <- read_excel("PSQ_Screener.xlsx")

```

### Get to know our data

Using the glimpse function shows that there are 46 variables in this dataset and 118 responses (this number will change throughout data collection). The questions of interest start at _1 and end at _18. When looking at the Perceived Stigma Questionnaire document, the Family Stigma Instrument (FAMSI), which is the real name of our questionnaire, indicates that there are several items with 5 factors. Data was not collected for Factor 1 since these are questions that 'take the perspective of someone not related to an individual with a disability.', Additionally, even though data was collected for Factor 5 (Q13-18), it will no be included into our analysis since they are related to positive aspects of caregiving and not stigma. Thus, the remaining questions to be analyzed which are part of three different factors are:

- Factor 2 (items 1-4): Affective affiliate stigma
- Factor 3 (items 5-8): Cognitive affiliate stigma
- Factor 4 (items 9-12): Behavioral affiliate stigma


```{r get to know our data}
# Use the glimpse function to see all of the variables in place
#glimpse(PSQ)

# Extract the variables of interest which are ID #, gender, and the items from above
PSQ2 <- PSQ %>%
  select(Child_Study_ID, Respondent_Gender, `_1_Ndilivwa_kuusa_ku_u_uujisi_buyumu_yumu`:`_18_Kubambilila_mwan_u_amikwasyi_beenzuma`)

# Rename the variables
names(PSQ2) <- c("ID", "Gender", paste("Q",1:18,sep=""))

# Remove QUestions from Factor 5
PSQ2 <- PSQ2 %>%
  select(-c(Q13, Q14, Q15, Q16, Q17))
```

### PSQ Screener (FAMSI) Quality control

```{r PSQ Screener Quality Control, results='asis'}
cat("1) There are",nrow(PSQ2),"responses to the FAMSI dataset.\n\n")
cat("2) There are",length(unique(PSQ2$ID)),"unique ID's in the FAMSI dataset.\n\n")
cat("3) The following ID's are duplicates that need to be corrected for:",paste(PSQ2$ID[duplicated(PSQ2$ID)], collapse = ", "),"\n\n")
cat("4) There are",sum(is.na(PSQ2)),"case(s) of missing data in this dataset.\n\n")
cat("5) In the meantime, any duplicate ID's and ID's with missing data will be dropped from further analysis until corrected for.\n\n")



PSQ2 %>%
  filter(!complete.cases(.)) %>%
  kbl(caption = " All ID's that are missing data") %>%
  kable_paper()


```

#### Dropping ID's with duplicate ID's code and/or missing data

```{r dropping duplicate IDs from further analysis}
# Drop any duplicate ID's
duplicate.IDs <- PSQ2$ID[duplicated(PSQ2$ID)]

# Remove them from the dataset
PSQ2 <- PSQ2 %>%
  filter(!(ID %in% duplicate.IDs))

# Remove any rows with missing data
PSQ2 <- PSQ2 %>%
  filter(complete.cases(.))
```


### Introducing DD Type from the screener information

Below we will load two datasets that differ from above. The first dataset contains some demographic information from our recruited sample plus their ID's, the second contains all of the information from the screener that we need- this does not contain ID's. We will join these datasets together to extract the types of DD our sample may have.


```{r load in the information from the IDs, warning = FALSE}
# Load in the information of children recruited into the study
setwd("~/KBB_new_2/matched_siblings")

# Load in their data
Recruited.Children <- read_excel("New ID Tracker.xlsx")

# Modify the variable 'x' to include DOB
Recruited.Children <- Recruited.Children %>%
  mutate(x = paste(x, Child_DOB))

# Load in the data of all screened children that have been scored
setwd("~/KBB_new_2/1_screener/final_data")

# Load in their data
All.Screened.Children <- read_excel("All Children.xlsx")

# Create the variable 'x' for All.Screened.Children so we can match them to Recruited.Children
All.Screened.Children <- All.Screened.Children %>% 
  mutate(x = paste(HOH_First_Name, HOH_Last_Name, Child_First_Name, Child_Last_Name, Child_Date_of_Birth))

# Extract the information that you need from Recruited Children
Recruited.Children2 <- select(Recruited.Children, ID, x) 

# Extract the information that you need from All Children
All.Screened.Children2 <- select(All.Screened.Children, KBB_DD_type, Child_age, KBB_DD_status, Child_Gender, epilepsy, x)

# Join the datasets together
joined.data <- Recruited.Children2 %>%
  left_join(All.Screened.Children2, by = "x")

```

#### Merge the data into one dataset


```{r what is the n for each group}
# Join this information with the data from our PSQ (regular)
PSQ2.final <- PSQ2 %>%
  left_join(joined.data, by = "ID")
```


### Creating Groups for DD Difficulties

We will be creating four groups, but are interested in analyzing the data from three. The four groups will be the following:

- Cognitive: These are children that only have cognitive related difficulties (e.g. Learning, Remembering, Communicating, etc.)
- Behavioral/Affective: These are children that have only behavioral or emotional difficulties (e.g. Self Care, Anxiety, Controlling Behavior, etc.)
- Both: Children that have difficulties for cognitive and behavioral/affective.
- None: Children that have no difficulties with CFM criteria. 

We split the children into groups to make the analysis easier to understand. Children that are in the none category will not be analyzed for they are not probands. There is an exception to this, there are some children who are proband that will be in the None group because they only have epilepsy. The table below will show each child with epilepsy status and how they fit into any of the four DD difficulty groups. 

```{r creating three groups}
# Create a template that can be used in the future to make modifications
Template <- c("CF14_Self_care", "CF15_Understood_Inside",  "CF16_Understood_Outside",  
              "CF17_Learning",  "CF18_Remembering", "CF20_Accepting_Challenge", 
              "CF21_Controlling_Behavior", "CF23_Anxiety", "CF24_Depression",          
              "CF22_Making_Friends", "CF19_Concentrating", "CF16_Controlling_Behavior")


# Create the requirements for the cognitive group
group1 <- c("CF15_Understood_Inside", "CF16_Understood_Outside", "CF17_Learning",
            "CF18_Remembering", "CF19_Concentrating")

# Create the requirements for the behavioral/affective group
group2 <- c("CF14_Self_care", "CF20_Accepting_Challenge", "CF21_Controlling_Behavior",
            "CF23_Anxiety", "CF24_Depression", "CF22_Making_Friends", "CF16_Controlling_Behavior")

# Collapse the groups with the or sign
group1.index <- paste(group1, collapse ="|")
group2.index <- paste(group2, collapse ="|")

# Introduce the cognitive group to the data
PSQ2.final <- PSQ2.final %>%
  mutate(cognitive.group = 
           case_when(
             grepl(group1.index, KBB_DD_type) ~ "Yes",
             TRUE ~ "No"
           )
         )

# Introduce the behavioral/affective group to the data
PSQ2.final <- PSQ2.final %>%
  mutate(behavioral.group = 
           case_when(
             grepl(group2.index, KBB_DD_type) ~ "Yes",
             TRUE ~ "No"
           )
         )

# Create a final variable called groups
PSQ2.final <- PSQ2.final %>%
  mutate(Group = 
           case_when(
             cognitive.group == "Yes" & behavioral.group == "Yes" ~ "Both",
             cognitive.group == "Yes" ~ "Cognitive",
             behavioral.group == "Yes" ~ "Behavioral/Affective",
             TRUE ~ "None"
             
           ))

# Quality Check
#PSQ2.final %>%
#  select(KBB_DD_type, KBB_DD_status, epilepsy, Group) %>%
#  view()

# Change the output of Epilepsy to make the table easier to read
PSQ2.final <- mutate(PSQ2.final, epilepsy = ifelse(epilepsy == "Yes",
                                                   "Yes Epilepsy",
                                                   "No Epilepsy"))

# Change the KBB_DD_status to a factor so you can make Yes appear first
PSQ2.final$KBB_DD_status <- factor(PSQ2.final$KBB_DD_status,
                                   levels = c("Yes", "No"),
                                   labels = c("DD/Proband", "noDD"))


# Save a table with epilepsy information + KBB DD status and the group they are in
epilepsy.table <- table(PSQ2.final$KBB_DD_status, PSQ2.final$Group, PSQ2.final$epilepsy)


# Extract information from the epilepsy table and save it as a dataframe
No.epilepsy.df <- data.frame(cbind(addmargins(epilepsy.table[ , ,"No Epilepsy"])))
Yes.epilepsy.df <- data.frame(cbind(addmargins(epilepsy.table[ , ,"Yes Epilepsy"])))

# Create a table with the epilepsy dataframes
cbind(Yes.epilepsy.df,
      No.epilepsy.df) %>%
  kbl(caption = "Proband difficulties type by epilepsy status") %>%
  kable_paper(full_width = F) %>%
  add_header_above(c(" " = 1, "Has Epilepsy" = 5, "Does not have Epilepsy" = 5)) %>%
  footnote("The table above shows the breakdown of the children in our sample. The first 4 columns show the number of children with epilepsy that are apart of the DD difficulty types groups. The next 4 columns are for the children with no epilepsy. This is important because some of the probands in our sample are defined as DD but do not have CFM DD status. Thus, the children that are DD/Proband but are in the None group will not be included in the analysis.")
```
### Keeping only probands

```{r keeping non overlapping groups}
# Create a dataset with only probands
PSQ2.final2 <- PSQ2.final %>%
  filter(Group %in% c("Cognitive", "Behavioral/Affective", "Both"))

# Obtain the table to see the sample sizes between these two groups
data.frame(n = cbind(addmargins(table(PSQ2.final2$Group)))) %>%
  kbl(caption = "Final sample size of our groups for analysis (Probands)") %>%
  kable_paper(full_width = F) %>%
  footnote("Sum indicates the addition of the sample sizes for the three groups. ")

```


### Obtaining the sum score for the Stigma Questionnaire

The responses from the Stigma Questionnaire (FAMSI) will be added into one sum score for each proband. As mentioned, the questions are asking caregivers how much perceived stigma they experience. Low experience to stigma result in a lower score and higher experience to stigma results in a higher score. Thus, the higher the sum score, the more stigma is experienced. 

```{r scoring the stigma questionnaire}
# Calculate the raw sum of the stigma questionnaire
PSQ2.final2$PSQ.Raw.Sum <- PSQ2.final2 %>%
  select(starts_with("Q")) %>%
  rowSums()

# Plot it between the groups
Cognitive.hist <- PSQ2.final2 %>%
  filter(Group == "Cognitive") %>%
  ggplot(aes(x= PSQ.Raw.Sum )) +
  geom_histogram(bins = 8,
                 fill = "white",
                 color = "black") +
  labs(title = "Cognitive PSQ Histogram") +
  scale_y_continuous(expand = c(0,0), limits = c(0,8)) + 
  theme_classic()

Behavioral.hist <- PSQ2.final2 %>%
  filter(Group == "Behavioral/Affective") %>%
  ggplot(aes(x= PSQ.Raw.Sum )) +
  geom_histogram(bins = 8,
                 fill = "white",
                 color = "black") +
  labs(title = "Behavioral/Affective PSQ Histogram") +
  scale_y_continuous(expand = c(0,0), limits = c(0,5)) + 
  theme_classic()

Both.hist <- PSQ2.final2 %>%
  filter(Group == "Both") %>%
  ggplot(aes(x= PSQ.Raw.Sum )) +
  geom_histogram(bins = 8,
                 fill = "white",
                 color = "black") +
  labs(title = "Cognitive and Behavioral/Affective PSQ Histogram") +
  scale_y_continuous(expand = c(0,0), limits = c(0,7)) + 
  theme_classic()


# Plot them
ggarrange(Cognitive.hist, Behavioral.hist,Both.hist)


```


### Run the Analysis

Three Mann-Whitney U Tests were run on the Perceived Stigma Sum Scores by Groups (Cognitive vs Behavioral/Affective vs Both).

'Wilcoxon W is simply the lowest sum of ranks but in order to calculate the p-value (Asymp. Sig), R uses an approximation to the standard normal distribution and also makes a continuity correction. The approximation is less reliable for small sample sizes'

This test is a pair-wise type test, meaning that each group will be compared to each other one at a time. If the perceived stigma sum scores are different between groups, then the returned p-value will be significant (p<.05).

```{r run the analysis, warning = FALSE}
# Create three pairs to do group comparisons
PSQ2.final2.Cognitive.Behavioral <-PSQ2.final2 %>%
  filter(Group != "Both")

PSQ2.final2.Both.Behavioral <-PSQ2.final2 %>%
  filter(Group != "Cognitive")

PSQ2.final2.Both.Cognitive <-PSQ2.final2 %>%
  filter(Group != "Behavioral/Affective")
         
    
# Run three Wilcox Tests
wilcox.test(PSQ2.final2.Cognitive.Behavioral$PSQ.Raw.Sum~PSQ2.final2.Cognitive.Behavioral$Group)
wilcox.test(PSQ2.final2.Both.Behavioral$PSQ.Raw.Sum~PSQ2.final2.Both.Behavioral$Group)
wilcox.test(PSQ2.final2.Both.Cognitive$PSQ.Raw.Sum~PSQ2.final2.Both.Cognitive$Group)
```
### Writing the results from the Mann-Whitney U Tests

There are significant group differences in perceived stigma between cognitive and behavioral/affective children (W = 49, p < .01). However, there were no significant differences in perceived stigma between behavioral/affective and both groups (W= 48, p = .05) and between cognition and behavioral/affective (W= 165, p = .76).


